|

### 2022-04-01 01-Chp-Background.Rmd
Range of corrections to Rmd files

Defined by the Diagnostic and Statistical Manual of Mental Disorders as a "major neurocognitive disorder", dementia is a progressive disease which impairs cognitive functions including speech, memory and executive reasoning.[@edition2013] At advanced stage, the condition causes severe behavioral and personality changes,[@cerejeira2012] culminating in reduced motor control that affects patients' ability to swallow or breathe.[@kumar2013] The condition has several distinct underlying causes, including Alzheimer's disease and vascular dementia.[@burns2009]

@@

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10-25% of [cases.]{.correction}[@alzheimersassociation2020] Vascular dementia is caused by a range of cerebrovascular disorders, and as a result, presentation  of symptoms can vary widely.[@iadecola2013] Similarly, due to the varied underlying  pathophysiology, vascular dementia can onset either quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts.[@venkat2015] Vascular dementia regularly co-occurs in patients with Alzheimer's disease.[@iadecola2013] This presentation is described as "mixed" dementia,[@custodio2017] and occurs in approximately 25% of cases.[@burns2009]

The remaining 10-30% of cases are caused by other dementia subtypes (e.g., Lewy body or frontotemporal dementias) or other neurological diseases (e.g., Parkinson's disease).[@burns2009]

@@

[However, evidence from longitudinal studies indicates the challenge of assigning a subtype to dementia outcomes, given the known heterogeneity in clinical phenotypes.]{.correction} [@rooney2017] [For example, ]{.correction} [@shim2013]

@@

Dementia is quickly becoming a critically important public health issue. Despite the age-specific incidence and prevalence of dementia remaining relatively constant over time,[@prince2016] an ageing population is set to create a dementia epidemic, particularly in westernised countries.[@flier2005] While approximately 525,000 patients have received a dementia diagnosis, the true number of people currently living with dementia in the UK is thought to be closer to 850,000, with this figure expected to double by 2040.[@baker2019] Globally, the prevalence of dementia is expected to reach 75 million by 2030.[@prince2016] Dementia is a leading cause of death in the UK,[@zotero-15757] and one of the few without a proven treatment.

Multimorbidity is common among people with dementia, with those diagnosed frequently having two or more other chronic conditions.[@poblador-plou2014]

Patients with dementia have a higher risk of diabetes, stroke, depression pneumonia, falling and infection

@@

Given the substantial burden that dementia represents and the absence of any curative therapies, as detailed in the above sections, the assessment of easily modifiable targets for their utility in the prevention of dementia should be prioritized.[@winblad2016] [Well-studied risk factors for the condition include cardiovascular factors (e.g. hypertension), education, diabetes, and smoking.]{.correction}[@anstey2019] The benefits of a prevention-based approach based on addressing known risk factors for dementia are well-studied. Reducing the prevalence of the seven most important risk factors for dementia (obesity, hypertension, diabetes, smoking, physical inactivity, [depression,]{.correction} and low educational attainment) by 10-20% per decade is estimated to result in a reduction in dementia prevalence of 8-15% by 2050.[@norton2014potential]

>>>>

Defined by the Diagnostic and Statistical Manual of Mental Disorders as a "major neurocognitive disorder", dementia is a progressive disease which impairs cognitive functions including speech, memory and executive reasoning.[@edition2013] At advanced stage, the condition causes severe behavioral and personality changes,[@cerejeira2012] culminating in reduced motor control that affects patients' ability to swallow or breathe.[@kumar2013]

The condition is thought to have several distinct underlying causes, the most common of which are Alzheimer's disease and vascular dementia.[@burns2009] Other causes include rarer dementia subtypes (e.g., Lewy body or frontotemporal dementias) and other neurological diseases (e.g., Parkinson's disease).[@burns2009]

@@

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10-25% of [cases.]{.correction}[@alzheimersassociation2020] Vascular dementia is caused by a range of cerebrovascular disorders, and as a result, presentation  of symptoms can vary widely.[@iadecola2013] Similarly, due to the varied underlying  pathophysiology, vascular dementia can onset either quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts.[@venkat2015]

[While widely used in the published literature, the validity of these dementia subtypes as distinct clinical outcomes is debatable.]{.correction} Vascular dementia regularly co-occurs in patients with Alzheimer's disease,[@iadecola2013] [a]{.correction} presentation described as "mixed" dementia.[@custodio2017] [Evidence from autopsies performed as part of longitudinal population based samples, such as the Medical Research Council Cognitive Function and Ageing Study,]{.correction}[@chadwick1992] [found that a majority of participants had evidence of both Alzheimer's disease and vascular pathology.]{.correction} [@neuropathologygroupofthemedicalresearchcouncilcognitivefunctionandageingstudymrccfas2001] [This suggests that the "pure" clinical subtypes of dementia are less common than the "mixed" outcome, particularly in older age groups,]{.correction}[@schneider2007] [and indicates that a composite all-cause dementia outcome might be considered more analytically valid than outcomes based on a single dementia subtype.]{.correction}

@@

@@

Dementia is quickly becoming a critically important public health issue. Despite the age-specific incidence and prevalence of dementia remaining relatively constant over time,[@prince2016] an ageing population is set to create a dementia epidemic, particularly in westernised countries.[@flier2005] While approximately 525,000 patients have received a dementia diagnosis, the true number of people currently living with dementia in the UK is thought to be closer to 850,000, with this figure expected to double by 2040.[@baker2019] Globally, the prevalence of dementia is expected to reach 75 million by 2030.[@prince2016] Dementia is a leading cause of death in the UK,[@zotero-15757] [and its impact on mortality is likely underestimated, given that death certificates frequently do not accurately capture dementia cases.]{.correction}[@ahmad2021; @gao2018] [Further,]{.correction} it is one of the few leading causes of death without a proven treatment.

[Additionally, multimorbidity is common among people with dementia, with those diagnosed frequently having two or more other chronic conditions,]{.correction}[@poblador-plou2014] [including diabetes, depression and cerebrovascular disease.]{.correction}[@bauer2014] [Dementia makes the diagnoses and management of comordid conditions harder, as people living with dementia may not be in a position to seek primary care]{.correction}[@cooper2017] [or self-manage treatments.]{.correction}[@sinclair2000]

@@

Given the substantial burden that dementia represents and the absence of any curative therapies, as detailed in the above sections, the assessment of easily modifiable targets for their utility in the prevention of dementia should be prioritized.[@winblad2016] [Previously-examined risk factors for the condition include cardiovascular factors (e.g. hypertension), education, diabetes, and smoking.]{.correction}[@anstey2019] The benefits of a prevention-based approach based on addressing known risk factors for dementia are well-studied. Reducing the prevalence of the seven most important risk factors for dementia (obesity, hypertension, diabetes, smoking, physical inactivity, [depression,]{.correction} and low educational attainment) by 10-20% per decade is estimated to result in a reduction in dementia prevalence of 8-15% by 2050.[@norton2014potential]

### 2022-04-01 05-Chp-CPRD-Analysis.Rmd
Range of corrections to Rmd files

Little evidence for an effect of lipid-regulating agents on the risk of Alzheimer's disease was found, with the exception of a slightly increased risk in those prescribed a certain type of lipid-regulating agent called fibrates. In contrast, I found an increased risk of vascular and other (i.e., non-Alzheimer's) dementia was associated with lipid-regulating agent use.

@@

* I found little evidence for an effect of lipid-regulating agents on probable or possible Alzheimer’s disease. However, lipid-regulating agent use was associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations for the "any LRA" analyses were driven by those observed in the large statin subgroup.

>>>>

[No]{.correction} evidence for an effect of lipid-regulating agents on the risk of Alzheimer's disease was found, with the exception of a slightly increased risk in those prescribed a certain type of lipid-regulating agent called fibrates. In contrast, I found an increased risk of vascular and other (i.e., non-Alzheimer's) dementia was associated with lipid-regulating agent use.

@@

* I found [very weak]{.correction} evidence for an effect of lipid-regulating agents on probable or possible Alzheimer’s disease. However, lipid-regulating agent use was associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations for the "any LRA" analyses were driven by those observed in the large statin subgroup.

### 2022-04-01 08-Discussion.Rmd
Added "and-meta-analyses" to chapter summary
Range of corrections to Rmd files

However, there are also four key limitations to this thesis. As with many studies of dementia outcomes, a limiting factor in the interpretation of the results is the absence of a detailed pathological mechanism. It is possible that statins have a true effect on dementia risk via some pathway other than lipid-regulation, but without knowledge of the mechanism, the plausibility of this relationship cannot be assessed.

@@

>>>>

However, there are also a number of key limitations to this thesis. As with many studies of dementia outcomes, a limiting factor in the interpretation of the results is the absence of a detailed pathological mechanism. It is possible that statins have a true effect on dementia risk via some pathway other than lipid-regulation, but without knowledge of the mechanism, the plausibility of this relationship cannot be assessed.

@@

[Fourthly, despite their wide use in the existing literature (as illustrated in the systematic review in Chapters]{.correction} \@ref(sys-rev-methods-heading) [&]{.correction} \@ref(sys-rev-results-heading) [), the clinical validity of dementia subtypes could be questioned and may lead to spurious results. This is supported by the difficulty in diagnosing a specific dementia subtype, and findings from population-based autopsies samples suggesting that "pure" Alzheimer's/vascular dementia outcomes are rare when compared to "mixed" dementia outcomes in older age groups, as discussed in detail in the Introduction (see Section]{.correction} \@ref(underlying-pathologies)[). In this case, results for the all-cause dementia outcome may be the most clinically useful, but were not the focus of the final chapters of these thesis.]{.correction}
